---
title: "Assignment 10B NobelPrize"
author: "Muhammad Ahmad"
format: html
editor: visual
---

### Approach 

My planned approach is to download the data from the API as a JSON file then 
covert into a tibble style data frame. Once I have the data frames I will answer
questions by cross referencing the two different data frames. The two data frames
are a Laurent data frame for information about the winners and a more simple df
about the awards.

## Retriving data from Nobel Prize API

The Nobel Prize offers two API endpoints. The laureates endpoint for information
about nobel prize winners and their personal information, and the prizes end 
point which returns general information about the prize category and who won it
during which year

```{r}
library(jsonlite)
library(tibble)
library(dplyr)

```

```{r}

laureates <- fromJSON(
  "https://api.nobelprize.org/2.1/laureates",
  flatten = TRUE
)$laureates


prizes_df <- fromJSON(
  "https://api.nobelprize.org/2.1/nobelPrizes",
  flatten = TRUE
)$nobelPrizes

 laureates <- as_tibble(laureates_df)
 prizes_df <- as_tibble(prizes_df)

 laureates_df_clean <- laureates_df %>%
   mutate(
     nobelPrizes = sapply(nobelPrizes, function(c){
       if(is.data.frame(c) ){
         paste(c,collapse = ",")
       } else {
         NA_character_
       }
       }))
     
 prizes_df_clean <- prizes_df %>%
   mutate(
     laureates = sapply(laureates, function(x) {
      if (is.null(x) || length(x) == 0) {
        NA_character_
      } else if (is.data.frame(x) && "knownName.en" %in% names(x)) {
        paste(x$knownName.en, collapse = ", ")       
      } else if (is.data.frame(x) && "fullName.en" %in% names(x)) {
        paste(x$fullName.en, collapse = ", ")      
      } else {
        NA_character_
      }
    })
  )

```

```{r}
head(laureates_df_clean)
```

```{r}
prizes_df_clean
```

### Questions 

## Question# 1 

What country has the most nobel prize winners that stay in their birth country?

```{r}
laureates_df_clean %>%
  group_by(birth.place.country.en) %>%
  summarise(
    laureates.born = n(),
    retained = sum(birth.place.countryNow.en == birth.place.country.en, na.rm = TRUE),
    ratio = retained / laureates.born
  )
```

This shows that the USA has the highest amount of nobel prize winners that stay
in their birth place.

## Question# 2

What age recived the most awards?


```{r}
laureates_df_clean$year.awarded <- substr(laureates_df_clean$nobelPrizes, 1, 4)

laureates_df_clean$age.awarded <- as.numeric(laureates_df_clean$year.awarded) -
  as.numeric(substr(laureates_df_clean$birth.date, 1, 4))

```

```{r}
laureates_df_clean %>%
  group_by(age.awarded) %>%
  summarise(
    total = n()
  )


```
Base off of this the age with the most awards is 53.
## Question# 3

On average which gender recieves the most awards?

```{r}
laureates_df_clean %>%
  filter(!is.na(gender)) %>%
  group_by(gender) %>%
  summarise(count = n())
```

here we can see our output shows us that according to the data frame more males 
Nobel prizes than women. 

## Question# 4
Which continent receives the most Nobel prizes? 
```{r}
laureates_df_clean %>%
  filter(!is.na(birth.place.continent.en)) %>% 
  group_by(birth.place.continent.en) %>%
  summarise(total_prizes = n()) %>%
  arrange(desc(total_prizes))   
```
Europe receives the most Nobel prizes according to the data.
