---
title: "Assignment 10B NobelPrize"
author: "Muhammad Ahmad"
format: html
editor: visual
---

### Approach 



### Challenges


## Retriving data from Nobel Prize API

The Nobel Prize offers two API endpoints. The laureates endpoint for information
about nobel prize winners and their personal information, and the prizes end 
point which returns general information about the prize category and who won it
during which year

```{r}
library(jsonlite)
library(tibble)
library(dplyr)

```

```{r}

laureates <- fromJSON(
  "https://api.nobelprize.org/2.1/laureates",
  flatten = TRUE
)$laureates


prizes_df <- fromJSON(
  "https://api.nobelprize.org/2.1/nobelPrizes",
  flatten = TRUE
)$nobelPrizes

 laureates <- as_tibble(laureates_df)
 prizes_df <- as_tibble(prizes_df)

 laureates_df_clean <- laureates_df %>%
   mutate(
     nobelPrizes = sapply(nobelPrizes, function(c){
       if(is.data.frame(c) ){
         paste(c,collapse = ",")
       } else {
         NA_character_
       }
       }))
     
 prizes_df_clean <- prizes_df %>%
   mutate(
     laureates = sapply(laureates, function(x) {
      if (is.null(x) || length(x) == 0) {
        NA_character_
      } else if (is.data.frame(x) && "knownName.en" %in% names(x)) {
        paste(x$knownName.en, collapse = ", ")       # use known name if available
      } else if (is.data.frame(x) && "fullName.en" %in% names(x)) {
        paste(x$fullName.en, collapse = ", ")         # fallback to full name
      } else {
        NA_character_
      }
    })
  )

 
laureates_df_clean %>%
  mutate(
    if (laureates_df_clean$knownName.en == prizes_df_clean$laureates){
        amountawarded = prizes_df_clean$prizeAmountAdjusted
        }else{
        NA_character_
        }
  )
```

```{r}
head(laureates_df_clean)
```

```{r}
prizes_df_clean
```

### Questions 

## Question# 1 

What country has the most nobel prize winners that stay in their birth country?

```{r}
laureates_df_clean %>%
  group_by(birth.place.country.en) %>%
  summarise(
    laureates.born = n(),
    retained = sum(birth.place.countryNow.en == birth.place.country.en, na.rm = TRUE),
    ratio = retained / laureates.born
  )
```

This shows that the USA has the highest amount of nobel prize winners that stay
in their birth place.

## Question# 2

What age recived the most awards?


```{r}
laureates_df_clean$year.awarded <- substr(laureates_df_clean$nobelPrizes, 1, 4)

laureates_df_clean$age.awarded <- as.numeric(laureates_df_clean$year.awarded) -
  as.numeric(substr(laureates_df_clean$birth.date, 1, 4))

```

```{r}
laureates_df_clean %>%
  group_by(age.awarded) %>%
  summarise(
    total = n()
  )


```
Base off of this the age with the most awards is 53.
## Question# 3

On average which gender recives the most awards?

```{r}


laureates_df_clean %>%
  filter(!is.na(gender)) %>%
  group_by(gender) %>%
  summarise(
    average.amount = mean(laureates_df_clean$prizeAmountAdjusted, na.rm = ),
    smallest = min(laureates_df_clean$prizeAmountAdjusted, na.rm = TRUE),
    largest = max(laureates_df_clean$prizeAmountAdjusted, na.rm = TRUE),
    total = n()
)
```

## Question# 4


### Conclusions 

