---
title: "Assignment 10A: Sentiment Analysis"
author: "Muhammad Ahmad"
format: html
editor: visual
---

## Approach

  After reading chapter 2 of *Text Mining with R* and understanding the code used for sentiment analysis written by Julia Silge and David Robinson, I believe I can adapt their example code to make estimation on sentiment for book reviews of the Brandon Sanderson series known as *Mistborn*. I am not certain on whether I want to use the lexicons in a package like tidy text or find one through my own research. I will deiced what my plan will be by the end of Friday but I may also experiment with different packages to see which lexicons will work with my text corpus.


## Challenges 

  Some challenges involved with this include fine tuning the sentiment analyse to my own goals, finding the appropriate lexicon, and being able to properly web scrape the reviews from a website. Spending extra time to experiment with the adapted code from the source to try and get a close expected outcome from the code. I assume that a lexicon with academic or review style language is best suited for this type of sentiment analyze and will look for that. I would like to web scrape a website for the reviews and put it in a data frame but if that takes more effort than I have time for then I will manually make a data frame. 


## Adapting the code 

```{r}
library(janeaustenr)
library(dplyr)
library(stringr)
library(tidyr)
library(ggplot2)
library(tidytext)
```


```{r}
get_sentiments("afinn")
get_sentiments("bing")
get_sentiments("nrc")
```

```{r}
library(janeaustenr)
library(dplyr)
library(stringr)

tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```

```{r}
nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice
#> # A tibble: 122,204 × 4
#>    book              linenumber chapter word     
#>    <fct>                  <int>   <int> <chr>    
#>  1 Pride & Prejudice          1       0 pride    
#>  2 Pride & Prejudice          1       0 and      
#>  3 Pride & Prejudice          1       0 prejudice
#>  4 Pride & Prejudice          3       0 by       
#>  5 Pride & Prejudice          3       0 jane     
#>  6 Pride & Prejudice          3       0 austen   
#>  7 Pride & Prejudice          7       1 chapter  
#>  8 Pride & Prejudice          7       1 1        
#>  9 Pride & Prejudice         10       1 it       
#> 10 Pride & Prejudice         10       1 is       
#> # ℹ 122,194 more rows
```

```{r}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

## Analysing Alice in Wonderland 

```{r}
library(syuzhet)
library(gutenbergr)
```
Shows that the id of Alice is 11
```{r}
gutenberg_works(title == "Alice's Adventures in Wonderland")
Alice_raw <- gutenberg_download(11)
```

```{r}
Alice_clean <- Alice_raw %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words)
```

```{r}
Alice_sentiment <- get_sentiment(Alice_clean$word, method = "syuzhet")

Alice_clean$sentiment_score <- Alice_sentiment
Alice_clean
```

```{r}
Alice_plot <- Alice_clean %>%
  mutate(index = row_number() %/% 200) %>%
  group_by(index) %>%
  summarise(sentiment = sum(sentiment_score))
```

```{r}
ggplot(Alice_plot, aes(index, sentiment)) +
  geom_col(fill = "seagreen") +
  theme_minimal() +
  labs(title = "Sentimental progression of Alice in Wonderland",
       subtitle = "Syuzhet Lexicon",
       x = "Narrative Progress ",
       y = "Total Sentiment Score")
```

```{r}
Alice_sentiment <- get_sentiment(Alice_clean$word, method = "afinn")

Alice_clean$sentiment_score <- Alice_sentiment
Alice_clean
```

```{r}
Alice_plot <- Alice_clean %>%
  mutate(index = row_number() %/% 200) %>%
  group_by(index) %>%
  summarise(sentiment = sum(sentiment_score))
```

```{r}
ggplot(Alice_plot, aes(index, sentiment)) +
  geom_col(fill = "seagreen") +
  theme_minimal() +
  labs(title = "Sentimental progression of Alice in Wonderland",
       subtitle = "afinn Lexicon",
       x = "Narrative Progress ",
       y = "Total Sentiment Score")
```


## Conclusions 
  I changed my initial plan to instead do a classic book hoping to make an more on par comparison. 
Group Alice by 200 words we see the sentimental progression of the story. There are many drops showing negative 
or low sentimental scores meaning dark theming happens but we get big pay offs of positivity towards the end after a large drop. 
The main difference between this and the example from chapter two is that this is a weighted total while the lexicons used in chapter 2
are sums. I also experiment with different lexicons to see if the graph would change at all. 
There was any strong changes the graph kept a similar shape except for slight changes in totals.

## Sources 

Silge, Julia, and David Robinson. Text Mining with R: A Tidy Approach. O'Reilly Media, 2017.


Code adapted from [Text Mining with R](https://www.tidytextmining.com)
by Julia Silge and David Robinson (O'Reilly, 2017),
licensed under CC BY-NC-SA 3.0.
